Deriving Multi-Headed Projective Dependency Parses \\ from Link Grammar Parses

### 2014-06-02 doc/LinksToDAG.tex
moved section on datasets
some clarifications etc.

For each undirected labeled edge $ij$ in the link corpus, where $i,j$ denote tokens of the same sentence with $i < j$, we introduce nonnegative integer variables $x_{ij}$ and $x_{ji}$ with a constraint $x_{ij}+x_{ji}=1$.  We interpret $x_{ij}=1$ or $x_{ji}=1$ to mean that the link has direction $i \rightarrow j$ or $i \leftarrow j$, respectively.\footnote{In practice we halve the number of variables by replacing $x_{ji}$ with $1-x_{ij}$ for $j > i$, but that obscures the exposition.}

For each non-0 token $v$, we ensure that it has at least one parent by constraining
\begin{align}\label{eqn:oneparent}
\sum_u x_{uv} & \geq 1
\end{align}
% OLD VERSION
% $$\sum_{u < v} x_{uv} + \sum_{u > v} \bar{x}_{vu} \geq 1$$
where $u$ ranges only over tokens such that the relevant variable exists.
%
To prevent cycles,\footnote{This also ensures {\sc reachable}, given \eqnref{oneparent}.} for each token $v$ we introduce a depth variable $d_v$ in the range $[0,n_v]$ (not constrained to be integer), where $n_v$ is the length of the sentence containing $v$.  We require a child's depth to be at least 1 greater than each of its parents' depths---since this can be satisfied iff the sentence has no directed cycles:
\begin{align}\label{eqn:nocycles}
(\forall u)\; d_v + (1+n_v) \cdot (1-x_{uv}) & \geq 1+d_u
% OLD VERSION
% $$(\forall u < v) d_v + (1+n_v) \cdot \bar{x}_{uv} \geq 1+d_u$$
% $$(\forall u > v) d_v + (1+n_v) \cdot x_{vu} \geq 1+d_u$$
\end{align}
The second summand ensures that \eqnref{nocycles} is trivially satisfied (hence has no effect) when $u$ is {\em not} the parent of $v$.  As a speedup, we can require the 0 tokens to have depth 0. \noteJE{does this really still make a difference to speed?  Possible better version: just add all the depths (perhaps multiplied by 0.001) to the objective function; this breaks ties in simplex.}

Finally, we encourage all links with the same label \noteJE{explain that we mean capital letters?} to have the same direction.  For each label $L$, we introduce binary variables $r_v$ and $\ell_v$ which say that a link of type $L$ is ``allowed'' to point right or left, respectively.  For each undirected edge $ij$ of label $L$, with $i < j$, we write
\begin{align}
x_{ij} &\leq r_L + s_{ij} & 
x_{ji} &\leq \ell_L + s_{ij}
\end{align}
where $s_{ij} \geq 0$ is a slack variable that allows an edge token to point in a disallowed direction if needed to ensure \eqnref{oneparent}--\eqnref{nocycles}.

@@

@@

@@

\NoteJE{how often is the ``both'' option used?}

\NoteJE{discuss differences!}

>>>>

@@

For each undirected labeled edge $ij$ in the link corpus, where $i,j$ denote tokens in the same sentence with $i < j$, we introduce nonnegative integer variables $x_{ij}$ and $x_{ji}$ with a constraint $x_{ij}+x_{ji}=1$.  We interpret $x_{ij}=1$ or $x_{ji}=1$ to mean that the link has direction $i \rightarrow j$ or $i \leftarrow j$, respectively.\footnote{In practice we halve the number of variables by replacing $x_{ji}$ with $1-x_{ij}$ for $j > i$, but that obscures the exposition.}

For each non-0 token $v$, we ensure that it has at least one parent by constraining\footnote{To denote two linked tokens, we use variables $i,j$ when $i$ is to the left of $j$, or variables $u,v$ when $u$ is the parent of $v$.}
\begin{align}\label{eqn:oneparent}
\sum_u x_{uv} & \geq 1
\end{align}
% OLD VERSION
% $$\sum_{u < v} x_{uv} + \sum_{u > v} \bar{x}_{vu} \geq 1$$
where $u$ ranges only over tokens such that the relevant variable exists.
%
To prevent cycles,\footnote{This also ensures {\sc reachable}, given \eqnref{oneparent}.} for each token $v$ we introduce a depth variable $d_v$ in the range $[0,n_v]$ (not constrained to be integer), where $n_v$ is the length of the sentence containing $v$.  We require a child's depth to be at least 1 greater than each of its parents' depths---constraints that can be satisfied iff the sentence has no directed cycles:
\begin{align}\label{eqn:nocycles}
(\forall u)\; d_v + (1+n_v) \cdot (1-x_{uv}) & \geq 1+d_u
% OLD VERSION
% $$(\forall u < v) d_v + (1+n_v) \cdot \bar{x}_{uv} \geq 1+d_u$$
% $$(\forall u > v) d_v + (1+n_v) \cdot x_{vu} \geq 1+d_u$$
\end{align}
The second summand ensures that \eqnref{nocycles} is trivially satisfied (hence has no effect) when $u$ is {\em not} the parent of $v$.  As a speedup, we can require the 0 tokens to have depth 0. \noteJE{does this really still make a difference to speed?  Possible better version: just add all the depths (perhaps multiplied by 0.001) to the objective function; this breaks ties in simplex.}

Finally, we encourage all links with the same label \noteJE{explain that we mean capital letters?} to have the same direction.  For each label $L$, we introduce binary variables $r_L$ and $\ell_L$, which say whether a link of type $L$ is ``allowed'' to point right or left, respectively.  For each undirected edge $ij$ of label $L$, with $i < j$, we write
\begin{align}
x_{ij} &\leq r_L + s_{ij} & 
x_{ji} &\leq \ell_L + s_{ij}
\end{align}
where $s_{ij} \geq 0$ is a slack variable that allows an edge token to point in a disallowed direction if needed to ensure \eqnref{oneparent}--\eqnref{nocycles}.

@@

\NoteJE{how often is the ``both'' option used?  and how much slack do we have?}

@@

\NoteJE{discuss linguistic differences!}
